MRD_final_microbiome

### 2019-08-04 GlobalMicrobiome_NS.Rmd
updated

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')
# test
#test
```

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')
# test
#test
#test
#clash 1
#clash2maryam

@@

###### INFO FOR PAPER - Specific to MND
```{r Eggerthella lenta}

# Eggerthella is a bacterial genus of Actinobacteria

# List of Genera in the Actinobacteria Phylum
taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Actinobacteria
print(taxa)

# Abundance of Eggerthella
head(abundances(pseq)["Eggerthella lenta et rel.",])

pseq.bac <- subset_taxa(pseq, Phylum == "Actinobacteria")

pseq.bac.g <- subset_taxa(pseq.bac, Genus == "Eggerthella lenta et rel.")

# Pick the taxa abundances for a given sample
# samplename <- sample_names(pseq)[,"Eggerthella lenta et rel."]

# Pick abundances for a particular taxon
# eg.abundances <- abundances(pseq, transform='compositional')[, samplename]

```{r Coprobacillus cateniformis}

# Coprobacillus cateniformis is a bacterium from the genus of Coprobacillus
# Coprobacillus is of phylum Firmicutes 
# Phylum = Firmicutes 
# Class = Erysipelotrichi

# This class of the Firmicutes Phylum isnt present in this data

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacilli
print(taxa)

```{r Parabacteroides goldsteinii and Parabacteroides distasonis}

# Parabacteroides goldsteinii 
# Phylum = Bacteriodetes

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
print(taxa)

# Parabacteroides distasonis et rel. is present
# RESEARCH DIFFERENCES

```{r Lactobacillus murinus and Lactobacillus gasseri}

# Lactobacillus murinus and gasseri
# Phylum = Firmicutes
# Class = Bacilli

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacilli
print(taxa)

# "Lactobacillus gasseri et rel."     
# "Lactobacillus plantarum et rel."  
#       -  closely related
# "Lactobacillus salivarius et rel." 
#       -  closely related according to https://microbiomejournal.biomedcentral.com/articles/10.1186/s40168-018-0440-5
# "Lactococcus"

```{r Prevotella melaninogenica}

# Prevotella melaninogenica
# Phylum = Bacteroidetes

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
print(taxa)

```{r Eisenbergiella tayi}

# Eisenbergiella tayi
# Phylum = Firmicutes
# Class = Clostridia
# Order = Clostridiales
# Family = Lachnospiraceae
# Produce butyric acid

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$
print(taxa)

# Not found

```{r Subdoligranulum variabile}

# Subdoligranulum variabile
# Phylum = Clostridium cluster IV
# Closely related to Faecalibacterium prausnitzii

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Clostridium_cluster_IV
print(taxa)

```{r Ruminococcus torques}

# Ruminococcus torques
# Phylum = Firmicutes
# Class = Clostridia
# Order = Clostridiales
# Family = Ruminococcaceae / Lachnospiraceae
# Genus = Ruminococcus

# Closely related to "Ruminococcus gnavus et rel."

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$
print(taxa)

```{r Akkermansia muciniphila}

# Akkermansia muciniphila
# Phylum = Verrucomicrobia
# Genus = Akkermansia

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Verrucomicrobia
print(taxa)

```{r Paper relative abundance}

paper <- subset_taxa(pseq, Phylum =="Actinobacteria" | Phylum =="Bacteroidetes"| Phylum =="Bacilli"  | Phylum =="Clostridium cluster IV"  | Phylum =="Verrucomicrobia")

p.f <- aggregate_taxa(paper, "Phylum")

p.f.rel <- microbiome::transform(p.f, "compositional")

plot.p.f.rel <- plot_composition(p.f.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

plot.p.f.rel + scale_fill_brewer(palette = "Paired")

plot(plot.p.f.rel + scale_fill_brewer(palette = "Paired") )

```{r Paper relative abundance}

paper <- subset_taxa(pseq, Genus =="Eggerthella lenta et rel." | Genus =="Parabacteroides distasonis et rel." | Genus =="Lactobacillus plantarum et rel."  | Genus =="Lactobacillus salivarius et rel." | Genus =="Prevotella melaninogenica et rel." | Genus =="Subdoligranulum variable at rel." | Genus =="Ruminococcus gnavus et rel." | Genus =="Akkermansia")

p.g.rel <- microbiome::transform(paper, "compositional")

plot.p.g.rel <- plot_composition(p.g.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

plot.p.g.rel + scale_fill_brewer(palette = "Paired")

plot(plot.p.g.rel + scale_fill_brewer(palette = "Paired") )

paper@tax_table

### 2019-08-04 MRD_final_microbiome .Rmd
updated

###### INFO FOR PAPER
```{r Eggerthella lenta}

@@

# Pick the taxa abundances for a given sample
samplename <- sample_names(pseq)[,"Eggerthella lenta et rel."]

# Pick abundances for a particular taxon
# eg.abundances <- abundances(pseq, transform='compositional')[, samplename]

@@

@@

@@

# Prevotella melaninogenica
# Phylum = Bacteroidetes

@@

@@

@@

@@

@@

@@

>>>>

###### INFO FOR PAPER - Specific to MND

```{r Taxa ranks in data set}

# Show available ranks in the dataset
rank_names(ps)

```{r Taxa ranks in data set}

```{r Violin plot}

# How many genera would be present after filtering?
length(get_taxa_unique(pseq, taxonomic.rank = "Genus"))
ps3 = tax_glom(pseq, "Genus", NArm = TRUE)

## VIOLIN PLOT EXAMPLE

# Defining a custom plot function, plot_abundance(), that uses phyloseq’s function to define a relative abundance graphic
# use this to compare more easily differences in scale and distribution of the abundance values in our phyloseq object before and after transformation
plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Eggerthella lenta et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

```{r Load packages}

library("knitr")
library("BiocStyle")

```{r Eggerthella lenta}

@@

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Eggerthella lenta et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Parabacteroides distasonis et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Lactobacillus gasseri et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

# Prevotella melaninogenica et rel.
# Phylum = Bacteroidetes

@@

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Prevotella melaninogenica et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Faecalibacterium prausnitzii et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$`Clostridium cluster XIVa`
print(taxa)

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Ruminococcus gnavus et rel."))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

plot_abundance = function(pseq,title = "",
                          Facet = "Genus", Color = "Phylum"){
  # Arbitrary subset, based on Phylum, for plotting
  p1f = subset_taxa(pseq, Genus %in% c("Akkermansia"))
  mphyseq = psmelt(p1f)
  mphyseq <- subset(mphyseq, Abundance > 0)
  ggplot(data = mphyseq, mapping = aes_string(x = "mnd",y = "Abundance",
                              color = Color, fill = Color)) +
    geom_violin(fill = NA) +
    geom_point(size = 1, alpha = 0.3,
               position = position_jitter(width = 0.3)) +
    facet_wrap(facets = Facet) + scale_y_log10()+
    theme(legend.position="none")
}

# Transform to relative abundance. Save as new object.
ps3ra = transform_sample_counts(ps3, function(x){x / sum(x)})

plotBefore = plot_abundance(ps3,"")
plotAfter = plot_abundance(ps3ra,"")
# Combine each plot into one graphic.
grid.arrange(nrow = 2,  plotBefore, plotAfter)

@@

paper@tax_table

@@

Note: Use of transformed/normalised and/or filtered datasets is optional however, it is recommended. IPCO internally check if there are zero abundance features and also check for common features (OTUs/genus/species) between reference taxonomy and queried dataset.
```{r IPCO}

# 1. Load reference data
install_github("https://github.com/IPCO-Rlibrary/IPCO")
library(ade4)

# 2. Normalise/Transform data
# Hellinger transformation
transform_data(pseq, relab)

# 3. Inferring functionality
# IPCO(R, L, Q)
# R dataset where rows are functions/pathways and columns are samples
# L dataset where rows are taxa/OTUs and columns are the same samples as dataset R
# Q dataset where rows are taxa/OTUs common between L and Q and columns are the samples for which functionality will be inferred

# Requires The references functional (R) and its paired taxonomy dataset (L) and the taxonomy dataset (Q) for which functions need to be inferred

IPCO(R, L, Q)

# 4. Checking for significant covariance between reference taxonomic and functional datasets 
```
